---
title: 'Module 1: Home Work'
author: "Lemi Daba"
date: "10/12/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Queston 12

Suppose that you want R to display ''Hello world!'' Drag and drop the phrases to create the correct R input. Note that there may be multiple ways to write this in R but we want you to use this command specifically.

### Answer
```{r Q_12}
print("Hello World!")
```

## Question 13

If you run the following code in R, what does the object my_sqrt contain?

```
z <- c(pi, 205, 149, -2)
y <- c(z, 555, z)
y <- 2 * y + 760
my_sqrt <- sqrt(y - 1)
```

### Answer

Let's run the code and observe the output.

```{r Q_13}
z <- c(pi, 205, 149, -2)
y <- c(z, 555, z)
y <- 2 * y + 760
my_sqrt <- sqrt(y - 1)

my_sqrt
```

It appears our output is a vector of length 9.

## Question 14

Assume that you tell R to divide zero by zero, what would you get?

### Answer

Let's try it!
```{r Q_14}
0/0
```
As can be observed, the result is `NaN` or Not a Number. `NaN` is one of four special values for doubles in R. The other three are `NA`, `Inf` and `-Inf`. 


## Question 15

If you have a missing value and you try to add it to a number, what result would you get?

### Answer

Let's give it a try!

```{r Q_15}
2 + NA
```
As you can see the answer is `NA`, a missing value. This is true in genera. Any arithmetic operation with a missing value returns a missing value.


## Question 16

We have asked the age of a group of 12 students. While 10 of them provided us with this information, 2 of them did not. We have constructed the vector age that captures this information.

```
age <- c(12, 28, 35, 27, NA, 25, 32, 45, 31, 23, NA, 34)
```

If we were interested in getting the vector without the missing values, which of the following lines of code would be useful to achieve this purpose? (Select all that apply)

### Answer

Let's locate (find the index position) of the missing values first. 

```{r include=FALSE}
age <- c(12, 28, 35, 27, NA, 25, 32, 45, 31, 23, NA, 34)
```

```{r}
which(is.na(age))
```

So the 5th and 11th entries are missing. Then all of the following are alternatives to extract only the non-missing values in `age`.

```{r}
age[-c(5, 11)]  # the - sign removes the 5th and 11th entries
```
```{r}
age[c(-5, -11)] # the - signs removes the 5th and 11th entries
```

```{r}
age[c(1, 2, 3, 4, 6, 7, 8, 9, 10, 12)]  # subsets based on the given index values
```

```{r}
age[!is.na(age)]  # uses logical subsetting to extract non-missing values (most efficient)
```


## Question 17

Download the data "CitesforSara.csv" into RStudio. This dataset includes paper-level citations from 1969 to 1998. First, read the CSV file into R using these commands:

```
library(tidyverse)
papers <- as_tibble(read_csv("[YOURFILEPATH]/CitesforSara.csv"))
```

Great! Let’s create a simplified dataset which only keeps the following variables contained in the papers dataset in this order: journal, year, cites, title, and au1. Use the method `select()` to accomplish this. Set this output to the variable `papers_select.` Drag and drop to create the code. Note that there may be more than one way to do this but there is only one correct answer from the following drag-and-drop options.


### Answer

Based on the drag and drop template:

```{r include=FALSE}
library(tidyverse)
papers <- read_csv("CitesforSara.csv")
```

```{r Q_17}
papers_select <- select(papers, journal, year, cites, title, au1)

# Let's see our data set:
papers_select
```


## Quesion 18

Let’s take a look at some of the most popular papers. Using the `filter()` method, how many records exist where there are greater than or equal to 100 citations?

### Answer

```{r Q_18}
papers_select %>%
  filter(cites >= 100)
```

As observed in the header, there are 205 records with at least 100 citations.


## Question 19

Use the `group_by()` function to group papers by journal. How many total citations exist for the journal “Econometrica”?

### Answer

```{r Q_19}
papers_select %>%
  group_by(journal) %>%
  summarise(tot_cites = sum(cites, na.rm = TRUE))
```

The journal Econometrica has 75789 total citations.


## Question 20

How many distinct primary authors (*au1*) exist in this dataset? Note: do not remove `NA` values.

### Answer

```{r Q_20}
papers_select %>%
  summarise(n_distinct(au1))
```

There are 2332 distinct primary authors.


## Question 21

Use the *dpylr* `contains()` method to create a new dataset called `papers_female` which contains only the columns from papers containing the string “female”. Drag and drop to create the code. Note that there may be more than one way to do this but there is only one correct answer from the following drag-and-drop options.

### Answer

Based on the drag-and-drop template:

```{r Q_21}
paper_female <- select(papers, contains("female"))

paper_female
```
























