---
title: 'Module 1: Home Work'
author: "Lemi Daba"
date: "10/12/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Queston 12

Suppose that you want R to display ''Hello world!'' Drag and drop the phrases to create the correct R input. Note that there may be multiple ways to write this in R but we want you to use this command specifically.

### Answer
```{r Q_12}
print("Hello World!")
```

## Question 13

If you run the following code in R, what does the object my_sqrt contain?

```
z <- c(pi, 205, 149, -2)
y <- c(z, 555, z)
y <- 2 * y + 760
my_sqrt <- sqrt(y - 1)
```

### Answer

Let's run the code and observe the output.

```{r Q_13}
z <- c(pi, 205, 149, -2)
y <- c(z, 555, z)
y <- 2 * y + 760
my_sqrt <- sqrt(y - 1)

my_sqrt
```

It appears our output is a vector of length 9.

## Question 14

Assume that you tell R to divide zero by zero, what would you get?

### Answer

Let's try it!
```{r Q_14}
0/0
```
As can be observed, the result is `NaN` or Not a Number. `NaN` is one of four special values for doubles in R. The other three are `NA`, `Inf` and `-Inf`. 


## Question 15

If you have a missing value and you try to add it to a number, what result would you get?

### Answer

Let's give it a try!

```{r Q_15}
2 + NA
```
As you can see the answer is `NA`, a missing value. This is true in genera. Any arithmetic operation with a missing value returns a missing value.


## Question 16

We have asked the age of a group of 12 students. While 10 of them provided us with this information, 2 of them did not. We have constructed the vector age that captures this information.

```
age <- c(12, 28, 35, 27, NA, 25, 32, 45, 31, 23, NA, 34)
```

If we were interested in getting the vector without the missing values, which of the following lines of code would be useful to achieve this purpose? (Select all that apply)

### Answer

Let's locate (find the index position) of the missing values first. 

```{r include=FALSE}
age <- c(12, 28, 35, 27, NA, 25, 32, 45, 31, 23, NA, 34)
```

```{r}
which(is.na(age))
```

So the 5th and 11th entries are missing. Then all of the following are alternatives to extract only the non-missing values in `age`.

```{r}
age[-c(5, 11)]  # the - sign removes the 5th and 11th entries
```
```{r}
age[c(-5, -11)] # the - signs removes the 5th and 11th entries
```

```{r}
age[c(1, 2, 3, 4, 6, 7, 8, 9, 10, 12)]  # subsets based on the given index values
```

```{r}
age[!is.na(age)]  # uses logical subsetting to extract non-missing values (most efficient)
```




