---
title: 'Module 2: Home Work'
author: "Lemi Daba"
date: "10/13/2020"
output: 
  github_document:    
    pandoc_args: --webtex 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preamble

For Questions 9 - 15, let $X \sim B(n, p)$ with $n = 8$ and $p = 0.2$.

For this section, please use this file for help with the R code. We highly encourage you to look up the documentation on your own time.

Look at the `rbinom()` documentation to sample draws from this binomial distribution. Use it to generate a vector called "successes" with 1000 draws from this distribution.

### Solution

```{r}
successes <- rbinom(n = 1000, size = 8, prob = 0.2)
```


## Question 9

Suppose you saved the output of previous step as "successes", which is a numeric vector.

Please fill in the blank to write down a simplest base code to plot a histogram of "successes":

___________(**successes**)


### Answer

```
hist(successes)
```


## Question 10

Which of the following histograms is closest to the plot that you created?

(The choices are not displayed here.)


### Answer

Running the previous code, we get

```{r}
hist(successes)
```


Looks like our result aligns most closely to the first choice.


## Question 12

R has two other built-in functions related to the binomial random variable. One is `dbinom()`, and the other is `pbinom()`. Look up these functions, and use them to answer the questions below.

For parts A through C, suppose you flip an unfair coin, where $ p(heads)=0.65 $. **Round your answer to two decimal places. For instance, if your answer is 0.467, please input 0.47**.

A. What is the probability of getting exactly 7 heads on 10 flips?

**Answer:**

```{r}
round(dbinom(x = 7, size = 10, prob = 0.65), 2)
```
B. What is the probability of getting at most 7 heads on 10 flips?

**Answer:**

```{r}
round(pbinom(q = 7, size = 10, prob = 0.65), 2)
```

C.What is the probability of getting at least 6 heads on 10 flips?

**Answer:**

```{r}
round(pbinom(q = 5, size = 10, prob = 0.65, lower.tail = FALSE), 2)
```

or 

```{r}
round(1 - pbinom(q = 5, size = 10, prob = 0.65), 2)
```

## Question 14


Admittedly, the histogram you generated in part 9 could use some work: Firstly, we plotted the frequency counts as opposed to the observed densities in our sample. Recall, that a probability function takes on values between 0 and 1. Second, it is not very pretty to look at.

We found the code below, to plot the densities from the “successes” vector you generated in part 9, however it is filled with blanks. Choose from the drop-down options for the correct code.

### Answer

Let's first load the tidyverse package.

```{r message=FALSE}
library(tidyverse)
```


```{r}
binom_draws <- as_tibble(data.frame(successes))

estimated_pf <- binom_draws %>%
  group_by(successes) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))

ggplot(estimated_pf, aes(x = successes, y = freq)) +
  geom_col() +
  ylab("Estimated Density")
```


## Question 15

Instead of plotting the observed density, we could’ve plotted the analytical densities derived from the formula above. Below is the code to do this. Select from the drop-down options the correct code.

Please note n = 1000 and p = 0.2

```
# Create a tibble with x and the analytical probability densities.
my_binom <- as_tibble(list(x=0:n, prob = dbinom(0:n, n, p)))
```

```
# Plot the computed theoretical density.
ggplot(my_binom, aes(x=x, y=prob)) + geom_col() +
     ylab("Analytical Density")
```

### Answer

Here, we create two vairables `n` and `p` and set their values to 1000 and 0.2, respectively.

```{r}
n <- 1000
p <- 0.2
```

Now, we just run the above code.

```{r}
my_binom <- as_tibble(list(x = 0:n, prob = dbinom(0:n, n, p)))

ggplot(my_binom, aes(x = x, y = prob)) +
  geom_col() +
  ylab("Analytical Density")
```


## Question 15 (continued)

Now, we are going to use the vector `my_binom` to compute the CDF. Look up `cummean` and `cumsum` and select the correct option to complete the code.

### Answer

```{r}
calculated_cdf <- my_binom %>%
  mutate(cdf = cumsum(prob))

# Plot the computed cdf
ggplot(calculated_cdf, aes(x = x, y = cdf)) +
  geom_step() +
  ylab("CDF")
```
